Field and experimental data together with computational models

### 2022-03-01 manuscript_1.0/manuscript_word.Rmd
Edited the manuscript according to Red1 comments. Readme.md file to facilitate reproducibility.

Often alternative cognitive mechanisms yield similar behaviour and/or cognitive
performances. This poses a problem for disentangling the mechanistic 
underpinnings of behaviour. This is particularly clear in research aimed at
discovering interspecific variation in *higher* cognitive abilities; or 
in other words, research on whether non-human animals show cognitive 
abilities believed to be uniquely human. For instance, when researchers 
try to find *mental time travel-like* behaviour, they usually come up with 
clever experiments showing the behaviour displayed requires inferences made 
through past events[@dally_Foodcaching_2006]. However, they often face challenges proposing alternative scenarios where simple associations can bring about the 
predicted behavioural outcome [@suddendorf_Evolution_2007]. Similarly, attempts 
to show *theory of mind* in non-human animals face objections 
justified by alternative mechanisms that bring about similar behavioural 
results [@heyes_Theory_1998]. Such controversies are usually settled by 
using the principle of parsimony, and it's cognitive version, 
Lloyd Morgan’s cannon. Which states that the simpler explanation (mechanism)
must be accepted. Ideally, however, alternative hypothesis should be evaluated 
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask the inter and 
intra-specific variation that arises in the interaction between mechanisms and 
environments. Thus, if predictions from the alternative mechanisms change
with environmental conditions, it is possible to asses the explanatory power 
of the different hypothesis by performing behavioural test in the context of
different environmental conditions.

From a evolutionary perspective, mechanisms are likely selected because 
of how they allow individuals to respond to environmental variation. 
For example, biological market theory predicts that the exchange rate of goods 
and/or services traded between cooperative partners adjusts to the 
law of supply and demand, when individuals have some degree of partner choice
[@noe_Biological_1995]. Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should favour the ability 
to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent 
mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station at their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally refered as the ephemeral reward task. 
Individuals are assesed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. This interspecific
variation in performance, and primates' known cognitive abilities $\color{red}{\text{ref}}$, shows that performane in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this inter-spefic variation suggest
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

Recent research on the cognitive tool kit needed to solve the ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates  [@wismer_Cuebased_2019] while others find 
it in the food [@pretot_Comparative_2021; @pretot_Comparing_2016]. 
However, identifying cues as salient is not the only challenge 
of the ephemeral reward task, as revealed by proximate learning models. 
Such models allow varying the cognitive tool kit and to evaluate 
which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value differently
one stimulus alone from compound versions of it has been termed 
configurational learning, or chunking, 
or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is in principle 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, however, the two 
mechanisms should be evaluated in light of how well 
they explain the available data. Note that different fields interested in
cognition and desicion making use different words to refer to negative 
reinforcers [@quinones_Reinforcement_2019;@sutton_Reinforcement_2018]. 
In here, for the sake of simplicity and clarity, we will use the word 
'penalty' to refer to this mechanism which includes a negative reinforcer.

Over the last decade, our team tested over a hundred of wild-caught 
cleaner fish in the exact same paradigm of  of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaners should not give priority to visitors any more.

The interplay between the underling cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether chaining, penalty, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under low
local cleaner-to-client ratio.

@@

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three models ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chai %>% ning ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

@@

panel.rsqr.all<-
  ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
                  aes(y=model,x=Rsqrd,fill=model))+
  stat_halfeye(alpha=0.5,point_size=3)+
  theme_classic()+xlim(-0.5,0.25)+
  scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
  name = "Model", labels = labelsPlot.Scen)+
  scale_y_discrete(labels=labelsPlot.Scen)+
  geom_vline(xintercept = 0,color="black",size=0.5)+
  theme(legend.position = 'none')
          # c(.32, .65)
        # ,legend.key.size = unit(0.22,'cm'),
        # axis.text.y = element_blank(),
        # legend.text = element_text(size=7))

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centred around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models` fit 
favours the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though, the peak of the three
$pseudo-R^2$ distributions was not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of black line in Fig. \ref{fig:post} f). This
shows that, accounting for variation in the parameter estimates, the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

(ref:pred) Observed and predicted probability of choosing a visitor. 
Left-hand side panel: colour contour shows the prediction from the 
learning model using the mode of the posterior distributions of parameters 
recovered by the statistical analysis. Dots show the frequency of visitor 
choices for the 12 reef sites, as well as the corresponding relative 
cleaner fish abundance (x axis) and frequency of visitors clients leaving 
the cleaning station (y axis). Right-hand side panels: Variation of the 
predicted probabilities  of choosing a visitor over a resident and their 
observed values for 12 locations. Circles show the mean prediction for 
each location from 100 samples taken from the posterior distribution. 
Thick and thin bars show the 66 and 95% credible interval, respectively, 
taken from those posterior samples. Squares show the predictions used for 
panel on the right-hand side. Colour coding denotes different reef sites/year 
of the data collection (see Empirical data section). Black line corresponds 
to a perfect match between observed and predicted probabilities. Upper 
panels (a and b) show predictions from a model including chaining and 
penalty; middle panels (c and d) from a  model with only chaining; lower 
panel (e and f) from a model with only penalty.

```{r pred, fig.cap='(ref:pred)',out.width='100%',fig.height=5}

@@

The main reason for chaining and penalty to give different
predictions is the way that cleaner fish relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models, they all predict an increase preference 
for visitors as the visitor probability increases  (Fig. \ref{fig:pred}).
In contrast, cleaner fish relative abundance has a different effect in the
model with only chaining, compared to the other two models,
full model and penalty. In the model with only chaining, only intermediate 
cleaner fish abundance triggers a preference for the visitor 
clients (Fig. \ref{fig:pred} c). In both models with penalty, 
both intermediate and low cleaner abundance triggered a preference 
for the visitor (Fig. \ref{fig:pred} a,e). Note, however, we calculated 
preferences shown in figure \ref{fig:pred} left panels by using only the mode 
of the posterior distributions, and by holding constant the balance between 
resident and visitors´s abundances. Panels on the right, show how close 
predictions are from the observed data, allowing the balance between 
client types to vary, and using a set of samples from the posterior 
distribution.

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, penalty, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the these fish apply 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients, and generalized this rule to the lab conditions 
once being familiar with the task.

While all three models captured well the positive relation between visitor 
leaving behaviour and cleaner fish performance in
the market task [@triki_Biological_2019], only the chaining mechanism 
predicted that cleaner fish performance in the task should be low in 
habitats with low cleaner-to-client ratios, regardless of the 
visitor leaving probability. In contrast, models including 
negative reward predicted the highest performance in the ephemeral 
reward task when relative cleaner fish abundance is low, particularly 
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low relative cleaner fish abundances mean the market has an excess of demand for
cleaning services. In the models, this translates to a cleaning station
that is frequently full. Thus, when a visitors leaves, it is likely that the
cleaner fish will have access to another client in the next step. Therefore, 
there will not be much difference in future reward between choosing 
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward 
on cleaner fish preference is the opposite, as in a busy cleaning station, to
that of chaining. cleaner fish will get more often the resident-visitor
state and will develop a preference for the visitor faster. 
At high cleaner fish abundances, the resident-visitor state
becomes so rare that neither mechanism is very efficient 
at generating a preference for visitors. When facing the resident 
visitor choice it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, the models suggest that chaining is the cognitive
mechanism that allows cleaner fish to adaptively adjust to their 
biological market ecological conditions.

>>>>

Often alternative cognitive mechanisms yield similar behaviour and/or cognitive
performances. This poses a problem for disentangling the mechanistic 
underpinnings of behaviour. This is particularly clear in research aimed at
discovering between species variation in *higher* cognitive abilities; or 
in other words, research on whether non-human animals show cognitive 
abilities believed to be uniquely human. For instance, when researchers 
try to find *mental time travel-like* behaviour, they usually come up with 
clever experiments to show the behaviour displayed requires inferences made 
through past events[@dally_Foodcaching_2006]. However, they often face the
challenge of alternative scenarios where simpler explanations, 
like classic associative learning, can bring about the 
observed behavioural outcome [@suddendorf_Evolution_2007]. Similarly, attempts 
to demonstrate the presence of *theory of mind* in non-human animals 
face objections justified by alternative mechanisms underpinning 
similar behavioural results [@heyes_Theory_1998]. 
Such controversies are usually settled by using the principle 
of parsimony and its cognitive version, 
Lloyd Morgan’s cannon, which states that the simpler explanation (mechanism)
should be accepted. Ideally, however, alternative hypotheses should be evaluated 
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask between and within 
species variation that arises among different environments. Thus, if 
predictions from the alternative mechanisms depend on environmental 
conditions, it is possible to assess the explanatory power 
of the different hypotheses by performing behavioural test in the context of
different environmental conditions.

From an evolutionary perspective, mechanisms are likely selected because 
of how they allow individuals to respond to environmental variation. 
For example, biological market theory predicts that the exchange rate of goods 
and/or services traded between cooperative partners adjusts to the 
law of supply and demand, when individuals have some degree of partner choice
[@noe_Biological_1995]. Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should favour the ability 
to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent 
mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station in their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering the same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish has eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally referred as the ephemeral reward task. 
Individuals are assessed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. 
This between variation in performance, and primates' known cognitive abilities
$\color{red}{\text{ref}}$, shows that performance in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

Recent research on the cognitive tool kit needed to solve the ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates  [@wismer_Cuebased_2019] while others find 
it in the food [@pretot_Comparative_2021; @pretot_Comparing_2016]. 
However, identifying cues as salient is not the only challenge 
of the ephemeral reward task, as revealed by proximate learning models. 
Such models allow varying the cognitive tool kit and evaluating 
which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish face not only resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value 
one stimulus differently alone from compound versions of it has been termed 
configurational learning, chunking, 
or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is, in principle, 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, the standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, however, the two 
mechanisms should be evaluated in light of how well 
they explain the available data. Note that different fields interested in
cognition and decision making use different words to refer to negative 
reinforcers [@quinones_Reinforcement_2019;@sutton_Reinforcement_2018]. 
Here, for the sake of simplicity and clarity, we will use the word 
'penalty' to refer to this mechanism which includes a negative reinforcer.

Over the last decade, our team tested over a hundred wild-caught 
cleaner fish in the exact same paradigm of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaner fish should not give priority to visitors anymore.

The interplay between the underlying cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether chaining, penalty, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show a low preference for visitors under low
local cleaner-to-client ratio.

@@

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three models ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chaining ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

@@

panel.rsqr.all<-
  ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
                  aes(y=model,x=Rsqrd,fill=model))+
  stat_halfeye(alpha=0.5,point_size=3)+
  theme_classic()+xlim(-0.5,0.25)+
  scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
  name = "", labels = labelsPlot.Scen)+
  ylab("")+
  scale_y_discrete(labels=labelsPlot.Scen)+
  geom_vline(xintercept = 0,color="black",size=0.5)+
  theme(legend.position = 'none')
          # c(.32, .65)
        # ,legend.key.size = unit(0.22,'cm'),
        # axis.text.y = element_blank(),
        # legend.text = element_text(size=7))

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centred around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models` fit 
favours the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though the peak of the three
$pseudo-R^2$ distributions were not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of the black line in Fig. \ref{fig:post} f). This
shows that accounting for variation in the parameter estimates the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

(ref:pred) Observed and predicted probability of choosing a visitor. 
Left-hand side panel: colour contour shows the prediction from the 
learning model using the mode of the posterior distributions of parameters 
recovered by the statistical analysis. Dots show the frequency of visitor 
choices for the 12 reef sites, as well as the corresponding relative 
cleaner fish abundance (x axis) and frequency of visitors clients leaving 
the cleaning station (y axis). Right-hand side panels: Variation of the 
predicted probabilities  of choosing a visitor over a resident and their 
observed values for 12 locations. Circles show the mean prediction for 
each location from 100 samples taken from the posterior distribution. 
Thick and thin bars show the 66 and 95% credible interval, respectively, 
taken from those posterior samples. Squares show the predictions used for the
panel on the right-hand side. Colour coding denotes different reef site/year 
of the data collection (see Empirical data section). The black line corresponds 
to a perfect match between observed and predicted probabilities. Upper 
panels (a and b) show predictions from a model including chaining and 
penalty; middle panels (c and d) from a  model with only chaining; lower 
panels (e and f) from a model with penalty only.

```{r pred, fig.cap='(ref:pred)',out.width='100%',fig.height=5,cache=FALSE}

@@

The main reason for chaining and penalty to give different
predictions is the way that cleaner fish relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models, they all predict an increase preference 
for visitors as the visitor probability increases  (Fig. \ref{fig:pred}).
In contrast, cleaner fish relative abundance has a different effect in the
model with only chaining, compared to the other two models,
full model and penalty. In the model with only chaining, only intermediate 
cleaner fish abundance triggers a preference for the visitor 
clients (Fig. \ref{fig:pred} c). In both models with penalty, 
both intermediate and low cleaner abundance triggered a preference 
for the visitor (Fig. \ref{fig:pred} a,e). Note, however, we calculated 
preferences shown in figure \ref{fig:pred} left panels by using only the mode 
of the posterior distributions, and by holding constant the balance between 
resident and visitors´ abundances. Panels on the right, show how close 
predictions are from the observed data, allowing the balance between 
client types to vary and using a set of samples from the posterior 
distribution.

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, penalty, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the rule these fish applied 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients and generalized this rule to the lab conditions 
once being familiar with the task.

While all three models captured well the positive relationship between visitor 
leaving behaviour and cleaner fish performance in
the market task [@triki_Biological_2019], only the chaining mechanism 
predicted that cleaner fish performance in the task should be low in 
habitats with low cleaner-to-client ratios, regardless of the 
visitor leaving probability. In contrast, models including 
negative reward predicted the highest performance in the ephemeral 
reward task when relative cleaner fish abundance is low, particularly 
together with a high probability of visitor leaving (Fig. \ref{fig:pred}).
Low relative cleaner fish abundances mean the market has an excess of demand for
cleaning services. In the models, this translates to a cleaning station
that is frequently full. Thus, when a visitor leaves, it is likely that the
cleaner fish will have access to another client in the next step. Therefore, 
there will not be much difference in future reward between choosing 
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward 
on cleaner fish preference is the opposite, as in a busy cleaning station, to
that of chaining. cleaner fish will get more often the resident-visitor
state and will develop a preference for the visitor faster. 
At high cleaner fish abundances, the resident-visitor state
becomes so rare that neither mechanism is very efficient 
at generating a preference for visitors. When facing the 
resident-visitor choice, it is still best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, the models suggest that chaining is the cognitive
mechanism that allows cleaner fish to adaptively adjust to their 
biological market ecological conditions.
